---
title: "Benthic Chlorophyll"
author: "Olivia Aguiar"
date: "7/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(Biostatistics)
library(readxl)
```

#making master data frame
```{r}
chloro <- read_xlsx('Benthic_Chlorophyll.xlsx') %>% filter(Chl_a != 'NA')
```

#separating out the sites
#I did some work that I've written over here- making individual excel sheets and removing NAs- to get this to line up well with the snail data I'm taking 2016 out of the SITE SHEETS ONLY- full data is still on the mastersheet and Benthic_Chlorophyll.xlsx
```{r}
chloro1 <- read_xlsx('LUM1_Benthic_Chlorophyll.xlsx')
chloro2 <- read_xlsx('LUM2_Benthic_Chlorophyll.xlsx')
chloro3 <- read_xlsx('LUM3_Benthic_Chlorophyll.xlsx')

LUM1 <- read_xlsx('Graph_LUM1.xlsx') %>% na.omit()
LUM2 <- read_xlsx('Graph_LUM2.xlsx') %>% na.omit()
LUM3 <- read_xlsx('Graph_LUM3.xlsx') %>% na.omit()
```

#looking to see if theres any overarching differences between seasons
```{r}
season_order <- c('Winter', 'Spring', 'Summer', 'Fall')

chloro1 %>% ggplot() +
  geom_boxplot(aes(x=factor(Season, levels = season_order), y=Average_Chla)) + xlab('Season') + ylab('Average Chlorophyll A (mg/cm)') + ggtitle("LUM1")
chloro1.aov.season <- aov(chloro1$Average_Chla ~ chloro1$Season)
summary(chloro1.aov.season)

chloro2 %>% ggplot() +
  geom_boxplot(aes(x=factor(Season, levels = season_order), y=Average_Chla)) + xlab('Season') + ylab('Average Chlorophyll A (mg/cm)') + ggtitle("LUM2")
chloro2.aov.season <- aov(chloro2$Average_Chla ~ chloro2$Season)
summary(chloro2.aov.season)

chloro3 %>% ggplot() +
  geom_boxplot(aes(x=factor(Season, levels = season_order), y=Average_Chla)) + xlab('Season') + ylab('Average Chlorophyll A (mg/cm)') + ggtitle("LUM3")
chloro3.aov.season <- aov(chloro3$Average_Chla ~ chloro3$Season)
summary(chloro3.aov.season)
```

#scatterplot for benthic chlorophyl during study months
```{r}
chloro1 %>% ggplot(aes(x=Study_month, y=Average_Chla)) + geom_point(aes(x=Study_month, y=Average_Chla)) + ggtitle('LUM1') + xlab('Study Month') + ylab('Average Chlorophyll A (mg/cm')

chloro2 %>% ggplot(aes(x=Study_month, y=Average_Chla)) + geom_point(aes(x=Study_month, y=Average_Chla)) + ggtitle('LUM2') + xlab('Study Month') + ylab('Average Chlorophyll A (mg/cm')

chloro3 %>% ggplot(aes(x=Study_month, y=Average_Chla)) + geom_point(aes(x=Study_month, y=Average_Chla)) + ggtitle('LUM3') + xlab('Study Month') + ylab('Average Chlorophyll A (mg/cm')
#all of the sites seem to have a general trend of increasing chlorphyll as the study progresses
```

#trying to plot density along benthic chlorophyll- starting with LUM1
#http://www.countbio.com/web_pages/left_object/R_for_biology/R_fundamentals/multiple_curves_R.html
```{r}
# define 2 data sets
Study_month <- chloro1$Study_month %>% na.omit()
Chlorophyll <- chloro1$Average_Chla %>% na.omit()
Lit_density <- LUM1$average %>% na.omit()

# plot the first curve by calling plot() function
# First curve is plotted
plot(Study_month, Chlorophyll, type="o", col="blue", pch="o", lty=1, ylim=c(0,150), ylab="y" )

# Add second curve to the same plot by calling points() and lines()
# Use symbol '*' for points.
points(Study_month, Lit_density, col="red", pch="*")
lines(Study_month, Lit_density, col="red",lty=2)

# Adding a legend inside box at the location (2,40) in graph coordinates.
# Note that the order of plots are maintained in the vectors of attributes.
legend(1,150,legend=c("Chlorophyll","Lit_density"), col=c("blue","red"), title = 'LUM1',
                                   pch=c("o","*"),lty=c(1,2), ncol=1)
```

#LUM2
```{r}
# define 2 data sets
Chlorophyll2 <- chloro2$Average_Chla %>% na.omit()
Lit_density2 <- LUM2$average %>% na.omit()

# plot the first curve by calling plot() function
# First curve is plotted
plot(Study_month, Chlorophyll2, type="o", col="blue", pch="o", lty=1, ylim=c(0,300), ylab="y" )

# Add second curve to the same plot by calling points() and lines()
# Use symbol '*' for points.
points(Study_month, Lit_density2, col="red", pch="*")
lines(Study_month, Lit_density2, col="red",lty=2)

# Adding a legend inside box at the location (2,40) in graph coordinates.
# Note that the order of plots are maintained in the vectors of attributes.
legend(1,300,legend=c("Chlorophyll","Lit_density"), col=c("blue","red"), title = 'LUM2',
                                   pch=c("o","*"),lty=c(1,2), ncol=1)
```

#LUM3
```{r}
# define 2 data sets
Chlorophyll3 <- chloro3$Average_Chla %>% na.omit()
Lit_density3 <- LUM3$average %>% na.omit()

# plot the first curve by calling plot() function
# First curve is plotted
plot(Study_month, Chlorophyll3, type="o", col="blue", pch="o", lty=1, ylim=c(0,350), ylab="y" )

# Add second curve to the same plot by calling points() and lines()
# Use symbol '*' for points.
points(Study_month, Lit_density3, col="red", pch="*")
lines(Study_month, Lit_density3, col="red",lty=2)

# Adding a legend inside box at the location (2,40) in graph coordinates.
# Note that the order of plots are maintained in the vectors of attributes.
legend(1,350,legend=c("Chlorophyll","Lit_density"), col=c("blue","red"), title = 'LUM3',
                                   pch=c("o","*"),lty=c(1,2), ncol=1)
```

#I feel like a lot of the variation here is getting lost because they are very different in terms of magnitude
```{r}
cor.test(Lit_density, Chlorophyll)

cor.test(Lit_density2, Chlorophyll2)

cor.test(Lit_density3, Chlorophyll3)
```

#going to try to make a new data frame with benthic chlorophyll, lit density, and study month
```{r}
combined1 <- data.frame(average_Chla = chloro1$Average_Chla,
                        average_lit = LUM1$average,
                        study_month = chloro1$Study_month)
plot(combined1$average_lit ~ combined1$average_Chla)
#doesn't seem to be anything going on here

combined2 <- data.frame(average_Chla = chloro2$Average_Chla,
                        average_lit = LUM2$average,
                        study_month = chloro2$Study_month)
plot(combined2$average_lit ~ combined2$average_Chla)
#again nothing pops out, this is more of a horizontal line though

combined3 <- data.frame(average_Chla = chloro3$Average_Chla,
                        average_lit = LUM3$average,
                        study_month = chloro3$Study_month)
plot(combined3$average_lit ~ combined3$average_Chla)
#no relationship can be seen here
```

#does benthic chlorophyll vary by site
```{r}
chloro.site.aov <- aov(chloro$Chl_a ~ chloro$Site)
summary(chloro.site.aov)

chloro %>% ggplot(aes(x=Site, y=Chl_a)) +
  geom_boxplot() + xlab('Site') + ylab('Littoraria Density')
```

