---
title: "Snail Visualization2"
author: "Olivia Aguiar"
date: "8/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(Biostatistics)
library(readxl)
library(plotrix)
library(lubridate)
library(formattable)
```

```{r}
snail <- read_xlsx('LUMSnails_Density_Master_Olivia_08.03.21.xlsx')
snail$year <- as.character(snail$year)
snail$season <- snail$month_name
snail$season <- gsub('Dec', 'Winter', snail$season)
snail$season <- gsub('Jan', 'Winter', snail$season)
snail$season <- gsub('Feb', 'Winter', snail$season)
snail$season <- gsub('Mar', 'Spring', snail$season)
snail$season <- gsub('Apr', 'Spring', snail$season)
snail$season <- gsub('May', 'Spring', snail$season)
snail$season <- gsub('Jun', 'Summer', snail$season)
snail$season <- gsub('Jul', 'Summer', snail$season)
snail$season <- gsub('Aug', 'Summer', snail$season)
snail$season <- gsub('Sep', 'Fall', snail$season)
snail$season <- gsub('Oct', 'Fall', snail$season)
snail$season <- gsub('Nov', 'Fall', snail$season)

snail1 <- snail %>% subset(site == 'LUM1')
snail2 <- snail %>% subset(site == 'LUM2')
snail3 <- snail %>% subset(site == 'LUM3')
```

#data frame for lit averages by study month
```{r}
#LUM1
average1 <- snail1 %>%
  group_by(Date) %>%
  summarise(
    Season = season,
    Month = month_name,
    Year = year,
    Mean = mean(Li_density_m2),
    SE = std.error(Li_density_m2)
  )

#LUM2
average2 <- snail2 %>%
  group_by(Date) %>%
  summarise(
    Season = season,
    Month = month_name,
    Year = year,
    Mean = mean(Li_density_m2),
    SE = std.error(Li_density_m2)
  )

#LUM3
average3 <- snail3 %>%
  group_by(Date) %>%
  summarise(
    Season = season,
    Month = month_name,
    Year = year,
    Mean = mean(Li_density_m2),
    SE = std.error(Li_density_m2)
  )
```

#monthly plots for each year
#LUM1
```{r}
month_order <- c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
average1 %>% subset(Year == '2017') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM1 2017 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average1 %>% subset(Year == '2018') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM1 2018 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average1 %>% subset(Year == '2019') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM1 2019 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average1 %>% subset(Year == '2020') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM1 2020 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

month_order2 <- c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun')
average1 %>% subset(Year == '2021') %>% ggplot(aes(x=factor(Month, levels = month_order2), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM1 2021 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')
```

#LUM2
```{r}
average2 %>% subset(Year == '2017') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM2 2017 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average2 %>% subset(Year == '2018') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM2 2018 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average2 %>% subset(Year == '2019') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM2 2019 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average2 %>% subset(Year == '2020') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM2 2020 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average2 %>% subset(Year == '2021') %>% ggplot(aes(x=factor(Month, levels = month_order2), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM2 2021 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')
```

#LUM3
```{r}
average3 %>% subset(Year == '2017') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM3 2017 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average3 %>% subset(Year == '2018') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM3 2018 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average3 %>% subset(Year == '2019') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM3 2019 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average3 %>% subset(Year == '2020') %>% ggplot(aes(x=factor(Month, levels = month_order), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM3 2020 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')

average3 %>% subset(Year == '2021') %>% ggplot(aes(x=factor(Month, levels = month_order2), y=Mean)) + geom_point(aes(x=factor(Month, levels = month_order), y=Mean)) + ggtitle('LUM3 2021 Densities') + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + xlab('Month') + ylab('Littoraria Density (m2)')
```

#boxplot of years
```{r}
snail1 %>% ggplot() + geom_boxplot(aes(x = year, y = Li_density_m2)) + xlab('Year') + ylab('Littoraria Density (m2)') + ggtitle('LUM1')

snail2 %>% ggplot() + geom_boxplot(aes(x = year, y = Li_density_m2)) + xlab('Year') + ylab('Littoraria Density (m2)') + ggtitle('LUM2')

snail3 %>% ggplot() + geom_boxplot(aes(x = year, y = Li_density_m2)) + xlab('Year') + ylab('Littoraria Density (m2)') + ggtitle('LUM3')
```

#boxplot of sites
```{r}
snail %>% ggplot() + geom_boxplot(aes(x = site, y = Li_density_m2)) + xlab('Site') + ylab('Littoraria Density (m2)')
```

#seasonal between sites
```{r}
snail1 %>% ggplot() + geom_boxplot(aes(x = season, y = Li_density_m2)) + xlab('Site') + ylab('Littoraria Density (m2)') + ggtitle('LUM1')

snail2 %>% ggplot() + geom_boxplot(aes(x = season, y = Li_density_m2)) + xlab('Site') + ylab('Littoraria Density (m2)') + ggtitle('LUM2')

snail3 %>% ggplot() + geom_boxplot(aes(x = season, y = Li_density_m2)) + xlab('Site') + ylab('Littoraria Density (m2)') + ggtitle('LUM3')
```

#monthly plot of entire study period
```{r}
average1$Date<-my(average1$Date,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))
average2$Date<-my(average2$Date,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))
average3$Date<-my(average3$Date,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))

average1 %>% ggplot(aes(x = Date, y = Mean)) + geom_point(aes(x = Date, y = Mean)) + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + theme_bw() + xlab('Date') + ylab('Littoraria Density (m2)') + ggtitle('LUM1')

average2 %>% ggplot(aes(x = Date, y = Mean)) + geom_point(aes(x = Date, y = Mean)) + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + theme_bw() + xlab('Date') + ylab('Littoraria Density (m2)') + ggtitle('LUM2')

average3 %>% ggplot(aes(x = Date, y = Mean)) + geom_point(aes(x = Date, y = Mean)) + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + theme_bw() + xlab('Date') + ylab('Littoraria Density (m2)') + ggtitle('LUM3')
```

#making graph of periwinkle density during the whole study period
```{r}
grouped <- snail %>%
  group_by(Date) %>%
  summarise(
    Mean = mean(Li_density_m2),
    SE = std.error(Li_density_m2),
    Year = year
  )
grouped$Date<-my(grouped$Date,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))

grouped %>% ggplot(aes(x = Date, y = Mean)) + geom_point(aes(x = Date, y = Mean)) + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE)) + theme_bw() + geom_smooth(method = 'lm') + xlab('Date') + ylab('Littoraria Density (m2)')
```

#making a graph of proportion of periwinkle vs nerites through study period
```{r}
snailnew <- snail %>% select(Date, site, Li_density_m2, Mel_density_m2, Nu_density_m2, allsnail_density_m2) %>% group_by(Date) %>% 
  summarise(
    mean_lit = mean(Li_density_m2),
    litSE = std.error(Li_density_m2),
    mean_mel = mean(Mel_density_m2),
    melSE = std.error(Mel_density_m2),
    mean_ner = mean(Nu_density_m2),
    nerSE = std.error(Nu_density_m2))

snailnew$total <- snailnew$mean_lit + snailnew$mean_ner

snailnew$proplit <- snailnew$mean_lit/snailnew$total
snailnew$propner <- snailnew$mean_ner/snailnew$total
snailnew$proplitSE <- sqrt(snailnew$proplit * (1 - snailnew$proplit) / 9 )
snailnew$propnerSE <- sqrt(snailnew$propner * (1 - snailnew$propner) / 9 )


snailnew$Date<-my(snailnew$Date,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))

snailnew %>% ggplot() + geom_col(aes(x = Date, y = proplit)) + geom_smooth(aes(x = Date, y = proplit)) + ylab('Percent') + ggtitle('Percent Littoraria')


snailnew %>% ggplot() + geom_col(aes(x = Date, y = propner)) + geom_smooth(aes(x = Date, y = propner)) + ylab('Percent') + ggtitle('Percent Nerites')

```

#graph of % lit and nerites but as points
```{r}
snailnew$proplit <- percent(snailnew$proplit)

snailnew %>% ggplot(aes(x = Date, y = proplit)) + geom_point(aes(x = Date, y = proplit)) + geom_errorbar(aes(ymin = proplit-proplitSE, ymax = proplit+proplitSE)) + geom_smooth(method = 'lm') + scale_y_continuous(breaks = c(0.25, 0.50, 0.75, 1.00), labels = c('25%', '50%', '75%', '100%')) + theme_bw() + ylab('Percent') + ggtitle('Percent L. irrorata')

snailnew %>% ggplot(aes(x = Date, y = propner)) + geom_point(aes(x = Date, y = propner)) + geom_errorbar(aes(ymin = propner-propnerSE, ymax = propner+propnerSE)) + geom_smooth(method = 'lm') + scale_y_continuous(breaks = c(0.25, 0.50, 0.75, 1.00), labels = c('25%', '50%', '75%', '100%')) + theme_bw() + ylab('Percent') + ggtitle('Percent N. usnea')
```

#quick calculations?
```{r}
grouped$Mean %>% subset(grouped$Year == '2017') %>% mean()
#average in 2017 was 115.8519

grouped$Mean %>% subset(grouped$Year == '2021') %>% mean()
#average in 2021 was 64.59259
#44% reduction in lit density from 2017 to 2021
```

