---
title: "Water_Salinity"
author: "Olivia Aguiar"
date: "7/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(Biostatistics)
library(plotrix)
library(lubridate)
```

```{r}
big <- read_xlsx('MC_Salinity_WaterTemp.xlsx', sheet = 'MC_Salinity') %>% na.omit()
new <- read_xlsx('MC_WaterDepth_Salinity_WaterTemp.xlsx', sheet = 'Salinity')
combined <- rbind(big, new)
combined$`Month/Year` <- format(as.Date(combined$`Date/Time`), "%m/%Y")
combined$`Month/Date` <- format(as.Date(combined$`Date/Time`), '%m/%d')
combined$Year <- format(as.Date(combined$`Date/Time`), '%Y')
combined$Month <- format(as.Date(combined$`Date/Time`), '%m')
combined$Date <- format(as.Date(combined$`Date/Time`), '%m/%d/%Y')
sal <- combined %>% subset(Year == '2017' | Year == '2018' | Year == '2019' | Year == '2020' | Year == '2021')
```

#yearly boxplot
```{r}
sal %>% ggplot() + geom_boxplot(aes(x = Year, y = Sal)) + ylab('Water Salinity')

year.aov <- aov(Sal ~ Year, sal)
summary(year.aov)
TukeyHSD(year.aov)
#all of the years are different?
```

#monthly boxplot
```{r}
#2017
sal %>% subset(Year == '2017') %>% ggplot() + geom_boxplot(aes(x = Month, y = Sal)) + ylab('Salinity') + ggtitle('Water Salinity in 2017')

#2018
sal %>% subset(Year == '2018') %>% ggplot() + geom_boxplot(aes(x = Month, y = Sal)) + ylab('Salinity') + ggtitle('Water Salinity in 2018')

#2019
sal %>% subset(Year == '2019') %>% ggplot() + geom_boxplot(aes(x = Month, y = Sal)) + ylab('Salinity') + ggtitle('Water Salinity in 2019')

#2020
sal %>% subset(Year == '2020') %>% ggplot() + geom_boxplot(aes(x = Month, y = Sal)) + ylab('Salinity') + ggtitle('Water Salinity in 2020')

#2021
sal %>% subset(Year == '2021') %>% ggplot() + geom_boxplot(aes(x = Month, y = Sal)) + ylab('Salinity') + ggtitle('Water Salinity in 2021')

```

#seasonal boxplot
```{r}
sal$Season <- sal$`Month/Year`
sal$Season <- gsub('01/2017', 'Winter', sal$Season) 
sal$Season <- gsub('02/2017', 'Winter', sal$Season) 
sal$Season <- gsub('03/2017', 'Spring', sal$Season) 
sal$Season <- gsub('04/2017', 'Spring', sal$Season) 
sal$Season <- gsub('05/2017', 'Spring', sal$Season) 
sal$Season <- gsub('06/2017', 'Summer', sal$Season) 
sal$Season <- gsub('07/2017', 'Summer', sal$Season) 
sal$Season <- gsub('08/2017', 'Summer', sal$Season) 
sal$Season <- gsub('09/2017', 'Fall', sal$Season) 
sal$Season <- gsub('10/2017', 'Fall', sal$Season) 
sal$Season <- gsub('11/2017', 'Fall', sal$Season) 
sal$Season <- gsub('12/2017', 'Winter', sal$Season)

sal$Season <- gsub('01/2018', 'Winter', sal$Season) 
sal$Season <- gsub('02/2018', 'Winter', sal$Season) 
sal$Season <- gsub('03/2018', 'Spring', sal$Season) 
sal$Season <- gsub('04/2018', 'Spring', sal$Season) 
sal$Season <- gsub('05/2018', 'Spring', sal$Season) 
sal$Season <- gsub('06/2018', 'Summer', sal$Season) 
sal$Season <- gsub('07/2018', 'Summer', sal$Season) 
sal$Season <- gsub('08/2018', 'Summer', sal$Season) 
sal$Season <- gsub('09/2018', 'Fall', sal$Season) 
sal$Season <- gsub('10/2018', 'Fall', sal$Season) 
sal$Season <- gsub('11/2018', 'Fall', sal$Season) 
sal$Season <- gsub('12/2018', 'Winter', sal$Season) 

sal$Season <- gsub('01/2019', 'Winter', sal$Season) 
sal$Season <- gsub('02/2019', 'Winter', sal$Season) 
sal$Season <- gsub('03/2019', 'Spring', sal$Season) 
sal$Season <- gsub('04/2019', 'Spring', sal$Season) 
sal$Season <- gsub('05/2019', 'Spring', sal$Season) 
sal$Season <- gsub('06/2019', 'Summer', sal$Season) 
sal$Season <- gsub('07/2019', 'Summer', sal$Season) 
sal$Season <- gsub('08/2019', 'Summer', sal$Season) 
sal$Season <- gsub('09/2019', 'Fall', sal$Season) 
sal$Season <- gsub('10/2019', 'Fall', sal$Season) 
sal$Season <- gsub('11/2019', 'Fall', sal$Season) 
sal$Season <- gsub('12/2019', 'Winter', sal$Season) 

sal$Season <- gsub('01/2020', 'Winter', sal$Season) 
sal$Season <- gsub('02/2020', 'Winter', sal$Season) 
sal$Season <- gsub('03/2020', 'Spring', sal$Season) 
sal$Season <- gsub('04/2020', 'Spring', sal$Season) 
sal$Season <- gsub('05/2020', 'Spring', sal$Season) 
sal$Season <- gsub('06/2020', 'Summer', sal$Season) 
sal$Season <- gsub('07/2020', 'Summer', sal$Season) 
sal$Season <- gsub('08/2020', 'Summer', sal$Season) 
sal$Season <- gsub('09/2020', 'Fall', sal$Season) 
sal$Season <- gsub('10/2020', 'Fall', sal$Season) 
sal$Season <- gsub('11/2020', 'Fall', sal$Season) 
sal$Season <- gsub('12/2020', 'Winter', sal$Season)

sal %>% ggplot() + geom_boxplot(aes(x = Season, y = Sal)) + ylab('Water Salinity')
```

#point plot for entire study period
```{r}
average <- sal %>% 
  group_by(`Month/Year`) %>% 
  summarise(mean = mean(Sal, na.rm = TRUE),
            SE = std.error(Sal))
average %>% ggplot(aes(x=`Month/Year`, y=mean)) + geom_point(aes(x=`Month/Year`, y=mean)) + geom_errorbar(aes(ymin = mean-SE, ymax = mean+SE)) + xlab('Date') + ylab('Water Salinity') + theme(axis.text.x = element_text(size = 7, angle = 45), axis.ticks = element_blank()) + scale_x_discrete(limits = c('01/2017', '02/2017', '03/2017', '04/2017', '05/2017', '06/2017', '07/2017', '08/2017', '09/2017', '10/2017', '11/2017', '12/2017', '01/2018', '02/2018', '03/2018', '04/2018', '05/2018', '06/2018', '07/2018', '08/2018', '09/2018', '10/2018', '11/2018', '12/2018', '01/2019', '02/2019', '03/2019', '04/2019', '05/2019', '06/2019', '07/2019', '08/2019', '09/2019', '10/2019', '11/2019', '12/2019', '01/2020', '02/2020', '03/2020', '04/2020', '05/2020', '06/2020', '07/2020', '08/2020', '09/2020', '10/2020', '11/2020', '12/2020', '01/2021', '02/2021', '03/2021', '04/2021', '05/2021', '06/2021'))
```

#Neritinia usnea threshold: <18
#Littoraria irrorata threshold: >2
#looking at days below lit threshold
```{r}
less <- sal %>% subset(Sal < 2)

litdays <- less %>% 
  group_by(`Date`) %>% 
  summarise(mean = mean(Sal),
            SE = std.error(Sal))

litmonth <- less %>% 
  group_by(`Month/Year`) %>% 
  summarise(mean = mean(Sal),
            SE = std.error(Sal))

litdays2 <- read_xlsx('litdays.xlsx')
sum(with(litdays2,Year == "2017"))
#33 days in 2017
#6/18-6/19
#6/27-7/12
#7/18-7/29
#9/6
sum(with(litdays2,Year == "2018"))
#7 days in 2018
#3/21
#4/20
#4/27-4/28
#6/4-6/6
sum(with(litdays2,Year == "2019"))
#51 days in 2019
#1/5-1/6
#1/9-1/10
#1/20
#1/30
#3/4-3/6
#3/26-3/27
#3/31
#4/20-4/22
#4/26-4/27
#5/31
#6/2-6/4
#6/9-6/11
#6/25-6/26
#6/29
#7/8-7/12
#8/6-8/13
#8/20-8/30
sum(with(litdays2,Year == "2020"))
#31 days in 2020
#2/14
#3/1
#3/26-3/27
#4/1
#5/5-5/7
#6/12-6/13
#6/26-7/16

litdays2$`Date`<-mdy(litdays2$`Date`,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))
#####this is for month/date/year!!!! get it in a date format!!!
```

#looking at above nerite threshold
```{r}
more <- sal %>% subset(Sal > 18)

nerdays <- more %>% 
  group_by(`Date`) %>% 
  summarise(mean = mean(Sal),
            SE = std.error(Sal))

nerdays2$`Date`<-mdy(nerdays2$`Date`,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))


nerdays2 <- read_xlsx('nerdays.xlsx')
sum(with(nerdays2,Year == "2017"))
#45 days in 2017
#1/1
#3/5-3/8
#3/24-3/25
#3/28-3/30
#4/2-4/3
#4/11-4/23
#4/29-4/30
#11/21-11/22
#11/27-12/8
#12/16-12/17
#12/26
#12/28
sum(with(nerdays2,Year == "2018"))
#43 days in 2018
#1/7-1/12
#1/20-1/23
#1/25-1/27
#2/3-2/11
#2/13-2/14
#2/20-2/28
#3/5
#3/27-3/29
#5/2
#5/4-5/6
#11/24
#12/8
sum(with(nerdays2,Year == "2019"))
#9 days in 2019
#2/20
#4/4
#4/18
#10/11
#10/15
#10/18-10/21
sum(with(nerdays2,Year == "2020"))
#34
#8/26-8/27
#10/9-10/10
#10/20-10/28
#11/7-11/13
#11/17
#11/19-11/29
#12/3
#12/16
```

#looking at entire study period
```{r}
#this is how to specify which x intervals to label- labels according to month names at those breaks
#d17 %>% ggplot() + geom_line(aes(x = `Month/Date`, y = Depth)) + theme(axis.text.x = element_text(angle = 45)) + scale_x_discrete(labels = monthlabel, breaks = c('01/01', '02/01', '03/01', '04/01', '05/01', '06/01', '07/01', '08/01', '09/01', '10/01', '11/01', '12/01')) + ggtitle('2017')

#need to make averaged calculations, ugh
yearlabel <- c('2000', '2001', '2002', '2003', '2004', '2005', '2006', '2007', '2008', '2009', '2010', '2011', '2012', '2013', '2014', '2015', '2016', '2017', '2018', '2019', '2020', '2021')
#can't figure it out- going to move on to just comparing years

combined %>% na.omit() %>% ggplot() + geom_boxplot(aes(x = Year, y = Sal))

#combined %>% group_by(Date) %>% ggplot() + geom_point(aes(x=Date, y=Sal))

combined.days <- combined %>% select(`Month/Year`,Sal, Year, Month) %>% group_by(`Month/Year`,Month) %>% summarise (mean_sal=mean(Sal,na.rm = TRUE))

combined.days$`Month/Year`<-my(combined.days$`Month/Year`,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))

comb<-combined.days %>% ggplot() + geom_point(aes(x=`Month/Year`, y=mean_sal,color=Month)) + geom_smooth(aes(x = `Month/Year`, y = mean_sal))

```

#looking at thresholds again but this time 2010-present
```{r}
litdaysnew <- combined %>% subset(Year == '2010' | Year == '2011' | Year == '2012' | Year == '2013' | Year == '2014' | Year == '2015' | Year == '2016' | Year == '2017' | Year == '2018' | Year == '2019' | Year == '2020' | Year == '2021') %>% subset(Sal < 2) %>% group_by(`Date`) %>% 
  summarise(mean = mean(Sal),
            SE = std.error(Sal))
litdaysnew$Date<-mdy(litdaysnew$Date, quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))

litdaysnew$Year <- format(as.Date(litdaysnew$Date), "%Y")

sum(with(litdaysnew,Year == "2010"))
#30
#1/5-1/13
#2/6-2/7
#2/10
#2/13
#2/16-2/19
#3/4
#3/13-3/18
#3/21-3/22
#3/28
#3/30
#6/18-6/19
sum(with(litdaysnew,Year == "2011"))
#7
#7/21
#7/24-7/25
#7/27
#8/3-8/5
sum(with(litdaysnew,Year == "2012"))
#0
sum(with(litdaysnew,Year == "2013"))
#0
sum(with(litdaysnew,Year == "2014"))
#9
#7/2
#7/16
#7/21-7/22
#7/26-7/30
sum(with(litdaysnew,Year == "2015"))
#16
#4/2
#4/20-4/21
#4/27
#4/30
#5/1-5/3
#6/11
#6/27-6/30
#7/1
#7/13-7/14
sum(with(litdaysnew,Year == "2016"))
#30
#1/23
#1/27-1/30
#2/9-2/13
#2/25-2/28
#3/23
#3/25
#4/5-4/8
#6/16
#6/18
#7/9
#8/14-8/15
#8/20-8/24
sum(with(litdaysnew,Year == "2021"))
#6
#4/21-4/22
#5/30
#6/4
#6/13-6/14
```

#now with nerites
```{r}
nerdaysnew <- combined %>% subset(Year == '2010' | Year == '2011' | Year == '2012' | Year == '2013' | Year == '2014' | Year == '2015' | Year == '2016' | Year == '2017' | Year == '2018' | Year == '2019' | Year == '2020' | Year == '2021') %>% subset(Sal > 18) %>% group_by(`Date`) %>% 
  summarise(mean = mean(Sal),
            SE = std.error(Sal))
nerdaysnew$Date<-mdy(nerdaysnew$Date, quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))

nerdaysnew$Year <- format(as.Date(nerdaysnew$Date), "%Y")

sum(with(nerdaysnew,Year == "2010"))
#5
#10/23-10/24
#10/26
#11/29
#12/29
sum(with(nerdaysnew,Year == "2011"))
#6
#1/9-1/10
#1/18-1/19
#1/24-1/25
sum(with(nerdaysnew,Year == "2012"))
#24
#10/23-10/28
#11/2-11/6
#11/9-11/11
#11/13-11/22
sum(with(nerdaysnew,Year == "2013"))
#1
#9/21
sum(with(nerdaysnew,Year == "2014"))
#55
#1/25
#2/4
#3/27-3/29
#4/3
#4/6
#4/14
#4/28
#5/8-5/14
#10/19-10/24
#10/28
#11/4-11/11
#11/17
#11/21-11/25
#12/5-12/11
#12/14-12/14
sum(with(nerdaysnew,Year == "2015"))
#24
#1/2
#1/11-1/12
#2/20-2/22
#2/25
#2/27-3/5
#3/7-3/13
#11/17-11/18
#12/13
sum(with(nerdaysnew,Year == "2016"))
#19
#10/15-10/16
#11/7
#11/28-11/30
#12/15-12/18
#12/23-12/31
sum(with(nerdaysnew,Year == "2021"))
#34
#1/1
#2/6-2/12
#2/14-2/15
#2/17-2/18
#2/21-2/25
#3/6-3/19
#3/23-3/25
```

#making graph of monthly minimums of whole study period
```{r}
combined.days2 <- combined %>% select(`Month/Year`,Sal, Year, Month) %>% group_by(`Month/Year`,Month) %>% summarise(min_sal=min(Sal,na.rm = TRUE))

combined.days2$`Month/Year`<-my(combined.days2$`Month/Year`,quiet=FALSE, tz=NULL, locale=Sys.getlocale("LC_TIME"))

comb<-combined.days2 %>% na.omit() %>% ggplot() + geom_point(aes(x=`Month/Year`, y=min_sal,color=Month)) + geom_smooth(aes(x = `Month/Year`, y = min_sal)) + xlab('Date') + ylab('Salinity') + ggtitle('Minimum Salinity')
```

#graph of salinity from 2011 to present
```{r}
graphsal <- combined %>% subset(Year == '2011'| Year == '2012' | Year == '2013' | Year == '2014' | Year == '2015' | Year == '2016' | Year == '2017' | Year == '2018' | Year == '2019' | Year == '2020' | Year == '2021') %>% select(`Month/Year`,Sal, Year, Month) %>% group_by(`Month/Year`,Month) %>% summarise(mean_sal=mean(Sal,na.rm = TRUE), SE = std.error(Sal, na.rm = TRUE))

graphsal$`Month/Year` <- my(graphsal$`Month/Year`, quiet = FALSE, tz = NULL, locale = Sys.getlocale("LC_TIME"))

graphsal %>% ggplot(aes(x = `Month/Year`, y = mean_sal)) + geom_point(aes(x = `Month/Year`, y = mean_sal)) + geom_errorbar(aes(ymin = mean_sal-SE, ymax = mean_sal+SE)) + geom_smooth(method = 'lm') + theme_bw() + xlab('Date') + ylab('Salinity')
```

#making bar chart of number of days with FW salinity
```{r}
year <- c('2011', '2012', '2013', '2014', '2015', '2016', '2017', '2018', '2019', '2020', '2021')
number <- c(7, 0, 0, 9, 16, 30, 33, 7, 51, 31, 6)

bar <- data.frame(year, number)

bar %>% ggplot() + geom_col(aes(x = year, y = number)) + theme_bw() + xlab('Year') + ylab('Number of Days Below 2 PSU')
```

